Supplementary Information

### 2021-02-10 doc/manuscript/MEE_manuscript.Rmd
updates re climate records

Monthly climate data for 1901-2019 were obtained from CRU v.4.04 [@harris_updated_2014; @harris_version_2020], and in a few cases corrected based on more local records (Appendix `r climate_correction`). 
Variables considered here included average daily minimum, maximum, and mean temperatures ($T_{min}$, $T_{max}$, $T_{mean}$, respectively); precipitation ($PPT$); and, when deemed reliable (Appendix `r climate_correction`), potential evapotranspiration ($PET$) and precipitation day frequency ($PDF$).
All ForestGEO climate records used here are archived in the ForestGEO Climate Data Portal, v1.0 [@anderson-teixeira_forestgeoclimate_2020].

@@

@@

# Acknowledgements {-#Acknowledgements}
Thanks to Helene Muller-Landau and Pete Kerby-Miller for bark thickness data.
Helpful feedback was provided by Helene Muller-Landau, Albert Kim...
This analysis was funded by a Smithsonian Scholarly Studies grant to KAT, SM, HCM, and CP.
The participation of PS, JK, and IV from the Czech Republic was supported by the Czech Science Foundation, project No. 19-09427S

>>>>

Monthly climate data for 1901-2019 were obtained from CRU v.4.04 [@harris_updated_2014; @harris_version_2020], and in a few cases corrected based on more local records (Appendix `r climate_correction`). 
Variables considered here included average daily minimum, maximum, and mean temperatures ($T_{min}$, $T_{max}$, $T_{mean}$, respectively); precipitation ($PPT$); and, when deemed reliable (Appendix `r climate_correction`), potential evapotranspiration ($PET$) and precipitation day frequency ($PDF$).
For the one riparian site, NIO, we tested for an effect of stream flow ($SF$), for which we obtained data for the Sparks, Nebraska station (station code: 06461500; 42°54'14"N, 100°26'13"W) from the U.S. Geological Survey (USGS) National Water Information System (https://waterdata.usgs.gov/nwis/uv/?site_no=06461500&agency_cd=USGS&referred_module=sw). 
All ForestGEO climate records used here are archived in the ForestGEO Climate Data Portal, v1.0 [@anderson-teixeira_forestgeoclimate_2020].

@@

*(predominantly negative temperature responses imply likely growth reductions under warming)*

@@

# Acknowledgements {-#Acknowledgements}
Thanks to Helene Muller-Landau and Pete Kerby-Miller for bark thickness data.
Helpful feedback was provided by Helene Muller-Landau, Albert Kim...
This analysis was funded by a Smithsonian Scholarly Studies grant to KAT, SM, HCM, and CP.
The participation of PS, JK, and IV from the Czech Republic was supported by the Czech Science Foundation, project No. 19-09427S.
PRISM climate data were purchased under an NSF grant to KAT (#DEB-1353301).

### 2021-02-10 doc/manuscript/MEE_manuscript_SI.Rmd
updates re climate records

*(documented in [GitHub issue #19 in ForestGEO_dendro](https://github.com/EcoClimLab/ForestGEO_dendro/issues/19))*

@@

$$DBH_Y  = DBH_{m} - 2*\left[r_{bark,m} - r_{bark,Y} + \sum_{y=Y}^{m} RW_{y} \right]$$
Here, $r_{bark}$ denotes bark thickness, which was estimated on a species-specific basis from allometries (Tables `r species_table`, `r bark_allometries`).
If $DBH_m$ was taken before coring, $RW$ values between $y=m$ and $y=Y$ were added instead of subtracted.
**If after...**

@@

*For BCNM, we calculated monthly $PPT$ and $PDF$ from daily precipitation readings made on BCNM starting in 1929 [@paton_barro_2019].*

>>>>

@@

$$DBH_Y  = DBH_{m} - 2*\left[r_{bark,m} - r_{bark,Y} + \sum_{y=Y}^{m} RW_{y} \right]$$
Here, $r_{bark}$ denotes bark thickness, which was estimated on a species-specific basis from allometries (Tables `r species_table`, `r bark_allometries`).
If $DBH_m$ was taken before coring, $RW$ values between $y=m$ and $y=Y$ were added instead of subtracted.
If $DBH_m$ was taken after coring, ***...(check code)**

@@

Monthly climate data for 1901-2019 were obtained from CRU v.4.04 [@harris_updated_2014; @harris_version_2020], and in a few cases corrected based on more local records. 
For BCNM, we calculated monthly $PPT$ and $PDF$ from daily precipitation readings made on Barro Colorado Island starting in 1929 [@paton_barro_2019] and used these to replace the corresponding CRU records.

Additionally, we used 800 m resolution PRISM records [@daly_mapping_2015; @daly_physiographically_2008-1] from 1930 to 2015 to test and correct for cases where the coarser-resolution CRU values were not locally representative. 
Variables in the PRISM data set included $T_{max}$, $T_{mean}$, $T_{min}$, and $PPT$.
To test for substantial differences between CRU and PRISM records, each variable was compared on a monthly basis using linear regression. We considered CRU records to be inappropriate if (1) a paired t-test revealed a significant difference between CRU and PRISM records, and (2) the absolute value of the mean monthly difference between CRU and PRISM exceeded 2.5°C (temperature variables) or 10 mm per month (precipitation).
Two sites met these criteria: LT and CB, both located in mountainous terrain in the western United States. 
For these sites, we used linear regression to characterize the relationship between the CRU and PRISM, and then used this relationship to correct CRU data outside the time frame of PRISM (i.e., pre-1930 and post-2015). 
CRU records were replaced with PRISM records over the time frame for which they were available.
For these sites, CRU records for variables not included in the PRISM data set ($PDF$, $PET$) were considered unreliable and excluded from the analysis.
